---
title: "Summative Assessment EMATM0061 - Part B"
author: "Akshat Pande | TX21857 | 2153363"
date: "<br> 12/01/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Section B

## B.1 Faulty Security System Probability

### Given
####Assumption - passage of time is per minute
$p^o$ - conditional probability that the sensor makes a sound if there is no person within one meter of the gate
<br>$p^1$ - conditional probability that the sensor makes at least once, if there is at least one person present
<br>$q$ - probability that at least one person walks within one meter of the gate
<br>$\phi$ - conditional probability that at one person has passed during the minute, and the alarm made the sound.

therefore,
$q^o$ - probability that nobody walks within one meter of the gate


Part (a):

Function used ->

$\phi$ = $(p^1/q)/(p^o/q^o + p^1/q)$


```{r phi function}

c_prob_person_given_alarm = function(p0,p1,q){
   
  (p1/q) / ((p0/(1-q)) + (p1/q))

}

```

Part (b):

Value of $\phi$ 
<br> Given: $p^o = 0.05,$ <br> $p^1=0.95$, <br> $q=0.1$

```{r testing phi function}

phi <- c_prob_person_given_alarm(0.05,0.95,0.1)

print(phi)

```


Part (c):

Graph between $\phi$ and $q$ for varying value of $q$
<br> Given : $p^o=0.05$ , <br> $p^1=0.95$

```{r graphplot}

library(ggplot2)

x1<-seq(from=0, to=1, by=.01)

print(x1)
plotphi<- as.data.frame(x1)
dim(plotphi)
x2<-seq(from=0, to=0.1, by=.01)
c_prob = function(p0,p1,input_vector){
   (p1/input_vector) / ((p0/(1-input_vector)) + (p1/input_vector))
}

x2 <-c_prob(0.05,0.95,x1)
x2[is.na(x2)]<-0
phivalue<-as.data.frame(x2)
plotphi$phivalue<-x2
plotphi <-setNames(plotphi,c("qvalue", "phivalue"))
head(plotphi)
flow<- ggplot(plotphi, aes(x=qvalue,y=phivalue) ) +geom_smooth()
flow
```




## B.2 Discrete function

### Given -
<br> $\alpha,\beta,\gamma \in [0,1]$
<br> X(DNV)- {0,1,2,5}
<br> P(X=1)= $\alpha$
<br> P(X=2)= $\beta$
<br> P(X=5)= $\gamma$
<br> P(X $\notin$ {0,1,2,5}) = 0


(a) Probability mass function = 
 
 P(x)=  {1 -$\alpha - \beta - \gamma$ , if x=0}<br>
        {$\alpha$, if x=1  }<br>
        {$\beta$, if x=2}<br>
        {$\gamma$, if x=5}<br>
        {0, otherwise}<br>
        
  
(b) E(X) = $\alpha + 2\beta + 5\gamma$

(c) E($X^2$) - $E(X)^2$ = $(\alpha + 4\beta + 25\gamma) - (\alpha + 2\beta + 5\gamma)^2$

<br>$X1,X2....,Xn$ <br>P(Xi=1)= $\alpha$, <br>P(Xi=2)= $\beta$, <br>P(Xi=5)= $\gamma$ and <br>P(X $\notin$ {0,1,2,5}) = 0

let $\tilde{X}$ = <br>$\frac{1}{n}$ $\sum_{i = 1}^{n}Xi$ be the sample mean

(d)  E(X) = $\frac{1}{n}$ $\sum_{i = 1}^{n}Xi$ * $(\alpha + 2\beta + 5\gamma)$ 

(e)  E($X^2$) - $E(X)^2$ = $\frac{1}{n}$ $\sum_{i = 1}^{n}Xi$ * ($(\alpha + 4\beta + 25\gamma) - (\alpha + 2\beta + 5\gamma)^2$)

(f) 
```{r function}
library('dplyr')
library('purrr')
sample_X_0125<-function(n,alpha,beta,gamma){
  sample_X<-data.frame(U=runif(n))%>%
  mutate(X=case_when(
  (0<=U)&(U<alpha)~1,
  (alpha<=U)&(U<alpha+beta)~2,
  (alpha+beta<=U)&(U<alpha+beta+gamma)~5,
  (alpha+beta+gamma<=U)&(U<=1)~0))%>%pull(X)
  return(sample_X)
}
```

(g)
```{r function2}

n<-100000
alpha<-1/10
beta<-2/10
gamma<-3/10

sample_X<-sample_X_0125(n,alpha,beta,gamma)
mean(sample_X)

```


(h)
```{r function3}

alpha<-1/10
beta<-2/10
gamma<-3/10
simulation_by_n<-data.frame(n=seq(0,10000,100))%>%
  mutate(sample_X=map(.x=n,~sample_X_0125(.x,alpha,beta,gamma)))%>%
  mutate(sample_avg=map_dbl(.x=sample_X,~mean(.x)))%>%
  select(-sample_X)%>%
  mutate(expectation=alpha+2*beta+5*gamma)
  simulation_by_n%>%head(5)

  
```


(i)
```{r function3plot}
plotit<- ggplot(simulation_by_n, aes(x=sample_avg)) + geom_histogram(binwidth=0.2)

plotit
```

(h)
```{r variance and expectation}
expectation<- alpha + 2*beta + 5*gamma
print(expectation)
var(sample_X)

```


## B.3 Exponential Distribution.

(a) Formula for population mean and variance

Population mean for exponential random variable X with paramater $\lambda$ = $\frac{1}{\lambda}$

Variance for exponential variable random variable X with parameter $\lambda$ = $\frac{2}{\lambda^2}$

(b) Formula for cumulative distribution function and quantile function for exponential random variables with paramter $\lambda$

Cummulative function is given by =  <br> $F\lambda(x)$ = {0, if x<=0}
                                                   {1 - $$e^{-\lambda x}$$, if x>0}

Quantile function is given by = <br> $$ F^{-1} \lambda (x)$$ = inf{ x $\in$ R : F $\lambda$ x <=p} 
                                                          {-$\infty$, if p=0}
                                                          { $\frac{1}{\lambda}$ ln(1-p), if p $\in$ (0,1]}
  
(c)

Maximum Likelihood Estimation = $\bar{X}$
  
                                                          
(d)

```{r function4}
library(dplyr)
set.seed(0)
num_trials_per_sample_size<-100
min_sample_size<-5
max_sample_size<-1000
sample_size_inc<-5
lambda_0<-0.01
poisson_simulation_df<-expand.grid(trial=seq(num_trials_per_sample_size),
sample_size=seq(min_sample_size,max_sample_size,
sample_size_inc))%>%
# create data frame of all pairs of sample_size and trial
mutate(simulation=pmap(.l=list(trial,sample_size),
.f=~rpois(.y,lambda=lambda_0)))%>%
# simulate sequences of Gaussian random variables
mutate(lambda_mle=map_dbl(.x=simulation,.f=mean))%>%
# compute the sample sd
group_by(sample_size)%>%
summarise(msq_error=mean((lambda_mle-lambda_0)*2))
poisson_simulation_df%>%
ggplot(aes(x=sample_size,y=msq_error))+
geom_smooth()+
theme_bw()+
xlab("Sample size")+ylab("Mean square error")
```


(e)

```{r bird_data}
library(readr)
bird_data<-read.csv("bird_data_EMATM0061.csv")

alpha<-0.00000004192201
sample_size<-length(bird_data$Time)
sample_mean<-mean(bird_data$Time)
sample_sd<-sd(bird_data$Time)
t<-qt(1-alpha/2,df=sample_size-1)
confidence_interval_l<-sample_mean-t*sample_sd/sqrt(sample_size)
confidence_interval_u<-sample_mean+t*sample_sd/sqrt(sample_size)
confidence_interval<-c(confidence_interval_l,confidence_interval_u)
print(confidence_interval)
percentage<-(confidence_interval[1] / confidence_interval[2]) * 100
print(percentage)
print("Confidence level of 95% achieved for alpha=0.00000004192201")
```