---
title: "Summative Assessment EMATM0061 - Part A"
author: "Akshat Pande | TX21857 | 2153363"
date: " <br> Date : 12/01/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section A

Data Wrangling tasks on financial data

## Code

Loading Libraries

```{r loading library, include=TRUE, echo=TRUE, message=FALSE}
library('readr') #tidyverse
library('utils')
library('psych')
library('plyr')
library('dplyr') #tidyverse
library('gmodels')
library('ggplot2') #tidyverse
library('doBy')
library('useful')
library('tidyr')
library('tools')
library('stringr')
```

## A.1 Reading CSV into Dataframe

### Tasks

- Downloading csv file
- Load csv file into R data frame
- Output the number of rows and column of the data frame

```{r dataframe, include=TRUE}
finance_data_original<-read_csv("finance_data_EMATM0061.csv", show_col_types = FALSE)
#Number of rows and column
dim(finance_data_original)

```

## A.2 Dataframe Column Manipulation

### Tasks

- New data frame (subset of original data frame)
- Corresponding columns for new data frame
- Display the first five rows and first three columns of the subset data frame

```{r column manipulation}
finance_data <- finance_data_original[c("state_year_code", "Details.Education.Education.Total","Details.Health.Health.Total.Expenditure","Details.Transportation.Highways.Highways.Total.Expenditure", "Totals.Revenue", "Totals.Expenditure")]
finance_data <- setNames(finance_data, c("state_year_code","education_expenditure","health_expenditure", "transport_expenditure", "totals_revenue", "totals_expenditure"))


head(finance_data[c(1:3)], n=5)
```

## A.3 Dataframe Column Analytical Addition

### Tasks

- Adding column to data frame
- Setting values to new column by using existing values of columns and applying calculations.
- Display the first three rows and four columns(as mentioned) of the new data frame

```{r column addition}

finance_data$totals_savings<-finance_data$totals_revenue - finance_data$totals_expenditure
dim(finance_data)
head(finance_data[c("state_year_code","totals_revenue","totals_expenditure","totals_savings")], n=3)
```


## A.4 Column Modification

### Tasks

- Separate column into multiple columns using given conditions
- Convert the column into specified formats and data types
- Display the first three rows and five columns(as mentioned) of the modified data frame

```{r column string manipulation and updation, warning=FALSE,}
finance_data<-finance_data %>% separate(state_year_code,
                c("state","year"), extra='drop')

finance_data$state<-str_to_title(finance_data$state)

head(finance_data[c(1,2)], n=3)
finance_data$year <- as.integer(finance_data$year)
head(finance_data[c("state","year","totals_revenue","totals_expenditure","totals_savings")], n=3)

```


## A.5 Plots

### Tasks

- Generate plot using given conditions
- Change attributes of the plot using given conditions
- Plot
```{r plottings, warning=FALSE, message=FALSE}
finance_data_plotL <- subset(finance_data, state == "Louisiana")
finance_data_plotM1 <- subset(finance_data, state == "Montana")
finance_data_plotM2 <- subset(finance_data, state == "Mississippi")
finance_data_plotK <- subset(finance_data, state == "Kentucky")
finance_data_plotL<-finance_data_plotL[c(1,2,6)]
finance_data_plotM1<-finance_data_plotM1[c(1,2,6)]
finance_data_plotM2<-finance_data_plotM2[c(1,2,6)]
finance_data_plotK<-finance_data_plotK[c(1,2,6)]


g<- ggplot(NULL, aes(x=year,y, lty=state, col=state, method="loess") ) +
geom_smooth(data=finance_data_plotL, mapping=aes(x=year,y=totals_revenue/1000000,)) +
geom_smooth(data=finance_data_plotM1, mapping=aes(x=year,y=totals_revenue/1000000)) +
geom_smooth(data=finance_data_plotM2, mapping=aes(x=year,y=totals_revenue/1000000)) +
geom_smooth(data=finance_data_plotK, mapping=aes(x=year,y=totals_revenue/1000000)) +
  scale_x_continuous(breaks = seq(0, 2020, 10)) +
  labs(title=" ", y="Revenue(millions $)", x="Year", caption="Source: finance_data")

g
```

## A.6 Get Decade Function

### Tasks

- Create function as specified
- Add column using the created function
- Show the three states with highest mean-average savings in decade 2000
```{r decade function}
retDecade = function(value){ return(value - (value%%10)) }
finance_data$decade<-retDecade(finance_data$year)

descstats<-function(x) {descStat(x, na.rm=TRUE)}

finance_data2000<-subset(finance_data, decade == 2000)

result<-summaryBy(totals_savings~state, data=finance_data2000, FUN=mean, na.rm=TRUE)
result$totals_savings.mean<-as.integer(result$totals_savings.mean)
result$decade<-2000
result<-result[order(-result$totals_savings.mean), ]  
head(result, n=3)
```

## A.7 State Summary

### Tasks

- Create Data frame for Alaska with decades
- Calculate mean and median for corresponding columns
- Display Alaska data frame

```{r alaska, warning=FALSE}
alaska_sum<-subset(finance_data, state=="Alaska")

alaska_summary<-summaryBy(education_expenditure~decade, data=alaska_sum, FUN=mean, na.rm=TRUE)
alaska_summary$ed_md<-summaryBy(education_expenditure~decade, data=alaska_sum, FUN=median, na.rm=TRUE)$education_expenditure.median
alaska_summary$he_mn<-summaryBy(health_expenditure~decade, data=alaska_sum, FUN=mean, na.rm=TRUE)$health_expenditure.mean
alaska_summary$he_md<-summaryBy(health_expenditure~decade, data=alaska_sum, FUN=median, na.rm=TRUE)$health_expenditure.median
alaska_summary$tr_mn<-summaryBy(transport_expenditure~decade, data=alaska_sum, FUN=mean, na.rm=TRUE)$transport_expenditure.mean
alaska_summary$tr_md<-summaryBy(transport_expenditure~decade, data=alaska_sum, FUN=median, na.rm=TRUE)$transport_expenditure.median
alaska_summary <- setNames(alaska_summary, c("decade","ed_mn","ed_md", "he_mn", "he_md", "tr_mn","tr_md"))
print(alaska_summary)

```

## A.8 Impute by median

### Tasks

- create impute_by_median function which takes vector numerical
- replace na values in the vector to median of the columns
- Create a data frame idaho_2000 for Idaho state in the decade 2000
- Apply impute_by_median function to idaho_2000
- Display a subset of idaho_2000_imputed

```{r function impute_median}

impute_by_median <- function(input_vector) {
  
   ifelse(is.na(input_vector),median(input_vector, na.rm=TRUE), input_vector)
  
}

idaho_2000<-subset(finance_data, state=="Idaho" & decade == "2000")
idaho_2000<-select(idaho_2000,-c(state,decade))

idaho_2000_imputed<-sapply(idaho_2000, impute_by_median)

idaho_2000_imputed<-as.data.frame(idaho_2000_imputed)

head(idaho_2000_imputed[c("year","health_expenditure","education_expenditure","totals_savings")], n=5)
```

## Including Plots


